\documentclass{article}
\usepackage[utf8]{inputenc}

\title{NPFL063 - HW 2}
\author{Thuong-Hai Pham}
\date{December 2017}

\usepackage{qtree}

\begin{document}

\maketitle

\section{Analyze the following sentences:}

\subsection{The monkey ate a banana and a big melon.}
\Tree [.S [.NP [.Det The ] [.N monkey ] ] [.V ate ] [.NP [.NP [.Det a ] [.N banana ] ] [.Conj and ] [.NP [.Det a ] [.Adj big ] [.N melon ] ] ] ]

\subsection{The small monkey gave me a banana.}
\Tree [.S [.NP [.Det The ] [.Adj small ] [.N monkey ] ] [.V gave ] [.NP [.Pron me ] ] [.NP [.Det a ] [.N banana ] ] ]

\subsection{The monkey gave a banana to me.}
\Tree [.S [.NP [.Det The ] [.N monkey ] ] [.V gave ] [.NP [.Det a ] [.N banana ] ] [.PP [.P to ] [.NP [.Pron me ] ] ] ]

\Tree [.S [.NP [.Det The ] [.N monkey ] ] [.V gave ] [.NP [.Det a ] [.N banana ] [.PP [.P to ] [.NP [.Pron me ] ] ] ] ]

\subsection{George saw the big dog of my friend.}
\Tree [.S [.NP [.N George ] ] [.V saw ] [.NP [.Det the ] [.Adj big ] [.N dog ] [.PP [.P of ] [.NP [.Det my ] [.N friend ] ] ] ] ]

\subsection{The shop is at the corner of the street.}
\Tree [.S [.NP [.Det The ] [.N shop ] ] [.V is ] [.PP [.P at ] [.NP [.Det the ] [.N corner ] [.PP [.P of ] [.NP [.Det the ] [.N street ] ] ] ] ] ]

\subsection{The parents of the groom and the bride.}
\Tree [.NP [.Det The ] [.N parents ] [.PP [.P of ] [.NP [.NP [.Det the ] [.N groom ] ] [.Conj and ] [.NP [.Det the ] [.N bride ] ] ] ] ]

\Tree [.NP [.NP [.Det The ] [.N parents ] [.PP [.P of ] [.NP [.Det the ] [.N groom ] ] ] ] [.Conj and ] [.NP [.Det the ] [.N bride ] ] ]

\subsection{His son sings and her daughter plays the piano.}
\Tree [.S [.S [.NP [.Det His ] [.N son ] ] [.V sings ] ] [.Conj and ] [.S [.NP [.Det her ] [.N daughter ] ] [.V plays ] [.NP [.Det the ] [.N piano ] ] ] ]


\section{}

\subsection{Two sentences that are correct according to both this grammar and grammar of
Standard English (GSE). Draw the corresponding syntactic trees.}

The boy sees a small book.

\Tree [.S [.NP [.Det The ] [.N boy ] ] [.V sees ] [.NP [.Det a ] [.Adj small ] [.N book ] ] ]

The girl gives a small book to the boy.

\Tree [.S [.NP [.Det The ] [.N girl ] ] [.V gives ] [.NP [.Det a ] [.Adj small ] [.N book ] ] [.PP [.P to ] [.NP [.Det the ] [.N boy ] ] ] ]

This sentence can also have a second parse tree as in question 1.c, which PP is attached to NP ``a small book" instead of S.

\Tree [.S [.NP [.Det The ] [.N girl ] ] [.V gives ] [.NP [.Det a ] [.Adj small ] [.N book ] [.PP [.P to ] [.NP [.Det the ] [.N boy ] ] ] ] ]

\subsection{A sentence that is correct according to this grammar, but incorrect according to
GSE. Where is the problem?}

``The boy sees a interesting book."

This sentence is correct according to the given grammar. However, `a' must be changed to `an' in GSE. Even if `an' is added to the `Det' category, the sentence above is still correct and one parser can not get these phenomenon right, yet which might be solved by a parser with probability.

\subsection{A sentence that is incorrect according to this grammar, but is correct according
to the GSE. Where is the problem?}

``The boy sees a small interesting book."
``The boy sees a man."

These two sentences are correct according to GSE but not the given grammar, which illustrate that the given grammar is too limited and can not cover the GSE vocabulary (`man' missing) or GSE's rule-set (two adjectives together: small interesting).

\end{document}
